Supplementary Material

### 2025-10-01 response.tex

  \begin{itemize}
  \item The writing is still conversational and the English is really poor and full of mistakes. For example,
  the authors use intensively the word “this” along the paper. Moreover, page 5 “the CDF and
  PDF are more used often for modeling and inference”; end of page 9, what does the sentence
  “But when these functions belong...” mean? Page 13, “to assess the how well...”; or “the the
  CDF ...”. On page 16 ,“decrase”

  \item Please include the formula along the text and not between dots. F.e. “the p-WD in terms of the
  corresponding QFs is defined in (15)”. should become “ the $p-WD$ in terms of the corresponding
  QFs is defined as...”. It is an example, but mainly all the equation needs to follow a similar
  specification.

  \end{itemize}

\section*{Reviewer 2}

I find the idea and motivation behind QGP compelling, and particularly 
attractive is that the general framework and the Stan code provided allow for 
flexible and relatively easy extension to other distributions beyond normal 
mixtures. The manuscript is coherent, clear, and well-structured. I believe 
there are some points that need further attention to enhance clarity.

\begin{itemize}

\item Clarification on the asymptotic nature of QGP:
 Since the QGP method relies on asymptotic results, I recommend explicitly 
 stating this in the abstract for clarity. Additionally, a concise summary in 
 the main text about when these asymptotic assumptions hold would be helpful. 
 While graphical results are provided, a practical guide on suitable sample 
 sizes or conditions would enhance the method's applicability.

\item Additional visual evaluation of estimated distributions:
 While thorough quantitative evaluations (KLD, Wasserstein) are provided, 
 additional graphical comparisons between the estimated and true underlying 
 distributions would offer valuable insight. After closely examining Figure 2, 
 it remains unclear to me why certain alternative methods underperform. For 
 instance, I suspect that the low coverage observed for the IND method is due 
 to its underestimation of uncertainty. Visualizing a set of estimated 
 distributions for selected scenarios (such as n=50 or n=150) would help reveal 
 where and why specific methods exhibit inaccuracies or weaknesses. These 
 visualizations would also support a more intuitive understanding of the 
 magnitude of approximation error; for example, it is not immediately obvious 
 whether a KLD of 0.05 is large or small.

\item Justification and selection of informative priors:
 Currently, the priors chosen for the parameters of the QGP method are generic 
 and identical across all examples, including the flu hospitalization case 
 study. Given the Bayesian nature of QGP, this aspect is critically important. 
 The thoughtful use of informative priors could substantially distinguish the 
 QGP model from competing methods (ORD, KDE, SPL). Specifically, incorporating 
 expert knowledge into the prior distributions, especially for practical 
 applications such as flu modeling, could lead to improved inference, 
 increased stability with smaller sample sizes, reduced uncertainty, and 
 better interpretability. It would strengthen the manuscript if the authors 
 explicitly discuss this point in the conclusions. In particular, it would be 
 valuable to explain how to select these priors, perform prior predictive 
 checks (see, for example, (1)), suggest context-specific informative priors, 
 and clarify how these choices could vary across different applications.
Reference:
(1) Betancourt, M. (2017). Principled Bayesian Workflow. Available at:
\href{https://betanalpha.github.io/assets/case_studies/principled_bayesian_workflow.html}{}

\item Sensitivity to the choice of distribution within QGP:
 While the paper fixes the QGP distributional assumption to a mixture of four 
 normal components, it would be informative to study how sensitive the results 
 are to this modeling choice. For example, it would be valuable to investigate 
 how varying the number of normal components in the mixture affects the 
 performance of QGP.

\item Explicit clarification on practical advantages of QGP:
 The manuscript carefully benchmarks QGP against alternative approaches, yet 
 the conclusions remain relatively modest regarding its practical advantages. 
 It would be helpful to include in the final discussion a concise summary of 
 the specific scenarios or data conditions where QGP outperforms the other 
 methods. Additionally, I believe that QGP stands out in two important aspects: 
 it naturally provides explicit uncertainty through the Bayesian posterior, and 
 it allows for the straightforward incorporation of expert knowledge via 
 informative priors.

\end{itemize}
\section*{Reviewer 3}

The paper introduces a “Quantile Gaussian Process” (QGP) model for estimating a 
full continuous distribution from a finite set of quantiles. The authors
apply this method in the context of probabilistic forecasting, particularly for
epidemic modeling, and compare the QGP approach with other quantile-matching
(QM) methods using both simulation studies and real CDC flu forecast data.

There are, however, several substantive issues with the manuscript, as outlined 
below. For these reasons, I do not believe it is currently suitable for
publication in Computational Statistics and Data Analysis. That said, I do not
wish to diminish the authors’ contributions to this important and active area
of research. With further development, this line of work could be expanded
into two complementary manuscripts: (i) a review-style paper summarizing 
existing quantile matching methods, and (ii) a computationally focused methods
paper with deeper theoretical or algorithmic contributions. And of course, that
will require substantial work beyond what is presented in this paper. The final
application section also has the potential to form the basis of a strong 
domainspecific paper, particularly in the biomedical or public health context. 
On a
personal note, I found the manuscript engaging, well-detailed and informative,
and I learned a great deal from reading it.

\begin{enumerate}[1]

  \item While the paper is well-written and the methodology is carefully 
  implemented, the contribution feels incremental rather than fundamentally
  novel. It reads more like an extended methodological review or a pedagogical 
  exposition rather than a research paper that breaks new ground.

  \item If the authors intend to emphasize their methodological contribution, 
  they
  should clearly articulate it in the introduction and present it more directly
  in Section 2. The focus should be placed primarily on describing the
  proposed method, rather than extensively reviewing existing literature.
  While such a review is valuable in a review article, in a research paper,
  much of this background could be moved to the supplementary material.

  \item For a stand-alone methods paper, the simulation study should be more
  comprehensive, incorporating additional tables and clearer, more informative 
  plots. The current figures make it difficult to assess the relative
  performance of the models.

\end{enumerate}

### 2025-10-06 supplemental.tex
remaking plots
change a couple things to supplement

%% The amssymb package provides various useful mathematical symbols
% \usepackage{amssymb}
%% The amsmath package provides various useful equation environments.
% \usepackage{amsmath}
\usepackage{amsmath,amsthm,amssymb,scrextend,mathtools}
\usepackage{hyperref}
\usepackage{newtxmath} %for making indicator function
% \RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\usepackage{natbib}

@@

Here we include a simulation study similar to the first study in section (??).
% \ref{sec:simulation_analyses}. 
However, instead of simulating data from a normal distribution, data is 
simulated 
from an exponential distribution with parameter $\lambda = 4$. The same prior 
distribution for $\lambda$ was used in fitting QGP, ORD, and IND models, and the 
same prior distribution was assigned to $n$ in QGP and ORD models and 
$1/\sigma_{\rho}$. The prior for $\lambda$ was 
$\pi(\lambda) \sim N(0, 7^2)\vmathbb{1}\{\lambda > 0\}$, the prior for $n$ was
$\pi(n) \sim N(0, 3000^2)\vmathbb{1}\{n > 0\}$, and the prior for
$1/\sigma_{\rho}$ was
$\pi(1/\sigma_{\rho}) \sim N(0, 3000^2)\vmathbb{1}\{1/\sigma_{\rho} > 0\}$.

Figure \ref{fig:exp_par_fit} shows examples of posterior sample density plots 
for $\lambda$ and $n$. QGP and ORD show similar distributions which tend to be 
a bit wider than the posterior distributions from the IND model. The left side 
of figure \ref{fig:exp_cov_dists} shows the percent coverage of the 500 
simulation replicates for $\lambda$ and $n$. Here, the IND model coverage is 
much closer to the nominal level than in the normal distribution case of the 
previous section. The right side of figure \ref{fig:exp_cov_dists} shows the 
average UWD1, TV, and KLD distances for the five QM methods. Again the 
parametric models tend to perform better than the non-parametric SPL and KDE 
methods with the SPL and KDE performing better as $n$ and $K$ increase.

\begin{figure}

\centering
% \begin{subfigure}
  \centering
  \hspace{-.8cm}
  \includegraphics[width=.94\linewidth]{Images/exponential_par_fit.png}
  % \caption{A subfigure}
  % \label{fig:sub1}
% \end{subfigure}%
% \begin{subfigure}
  \centering
  \hspace{-.9cm}
  \includegraphics[width=1.05\linewidth]{Images/exponential_n_fit.png}
  % \caption{A subfigure}
  % \label{fig:sub2}
% \end{subfigure}
\caption{Density plots of posterior distribution samples for the exponential 
parameters by QM for QGP, ORD, and IND models. QM was done on estimated 
quantiles from a exponential distribution with parameter $\lambda = 4$. The 
posterior densities are for $\lambda$ (left) and sample size $n$ (right). Plots 
are faceted by true sample size $n \in \{50, 150, 500, 1{,}000\}$ ($y$-axis) 
and number of quantiles $K \in \{7, 15, 23\}$ ($x$-axis). Vertical lines 
(black) show the value of the true parameter.}
\label{fig:exp_par_fit}
\end{figure}

\begin{figure}[hbt!]
% \begin{subfigure}
  \includegraphics[width=\linewidth]{Images/exponential_coverage.png}
  % \caption{}
  % \label{MLEDdet}
% \end{subfigure}\hfill % <-- "\hfill"
% \begin{subfigure}
  \includegraphics[width=\linewidth]{Images/exponential_dists.png}
  % \caption{}
  % \label{energydetPSK}
% \end{subfigure}
\caption{Posterior coverage (left) calculated as the percentage of times the 
true parameter fell within the modeled 90\% credible interval over the 500 
replications. Coverage is faceted by the exponential parameter 
$\lambda$ and $n$ with $K \in \{3, 7, 15, 23, 50\}$, and by increasing sample 
size ($x$-axis). The five models QGP, ORD, QGPN, ORDN, and IND are colored as 
shown the legend. The horizontal line (black) is at the nominal 90\% level. 
Only QGP and ORD appear for the parameter $n$ as they are the only two which 
estimate an unknown $n$.
Distance between the true distribution and the estimated QM predictive 
distribution (right) averaged over the 500 replications. Distances include 
UWD1, TV, and KLD for $K \in \{3, 7, 15, 23, 50\}$, and by increasing sample 
size ($x$-axis).}
\label{fig:exp_cov_dists}
\end{figure}

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.49\linewidth]{Images/quants_lp.png}
  % \caption{A subfigure}
  % \label{fig:sub1}
%\end{subfigure}%
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.49\linewidth]{Images/dens_lp.png}
  % \caption{A subfigure}
  % \label{fig:sub2}
%\end{subfigure}
\caption{QM fits of $K=23$ quantiles by KDE, SPL, IND, and QGP for $n \in \{50, 150, 500, 1{,}000\}$. The quantiles were sampled from the Laplace distribution $La(0,1)$. The quantile fits (left) show the true quantiles (black) with either the QM fit line (grey) or the credible intervals of 50\% (red) and 95\% (pink). 
The estimated PDF plots (right) show the true PDF (black) with either a the QM estimated PDF (grey) or the credible intervals of 50\% (red) and 95\% (pink).}
\label{fig:lp_fits}
\end{figure}

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.49\linewidth]{Images/quants_gmix.png}
  % \caption{A subfigure}
  % \label{fig:sub1}
%\end{subfigure}%
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.49\linewidth]{Images/dens_gmix.png}
  % \caption{A subfigure}
  % \label{fig:sub2}
%\end{subfigure}
\caption{QM fits of $K=23$ quantiles by KDE, SPL, IND, and QGP for $n \in \{50, 150, 500, 1{,}000\}$. The quantiles were sampled from the two component normal mixture distribution $w N(-1, 0.9) + (1-w)N(1.2, .6)$ where $w = 0.35$. The quantile fits (left) show the true quantiles (black) with either the QM fit line (grey) or the credible intervals of 50\% (red) and 95\% (pink). 
The estimated PDF plots (right) show the true PDF (black) with either a the QM estimated PDF (grey) or the credible intervals of 50\% (red) and 95\% (pink). }
\label{fig:gmix_fits}
\end{figure}

@@

% \begin{thebibliography}{00}
% 
% %% For authoryear reference style
% %% \bibitem[Author(year)]{label}
% %% Text of bibliographic item
% 
% \bibitem[Lamport(1994)]{lamport94}
%   Leslie Lamport,
%   \textit{\LaTeX: a document preparation system},
%   Addison Wesley, Massachusetts,
%   2nd edition,
%   1994.
% 
% \end{thebibliography}
\end{document}

>>>>

%% The amssymb package provides various useful mathematical symbols
% \usepackage{amssymb}
%% The amsmath package provides various useful equation environments.
% \usepackage{amsmath}
\usepackage{amsmath,amsthm,amssymb,scrextend,mathtools}
\usepackage{hyperref}
\usepackage{newtxmath} %for making indicator function
\usepackage{booktabs}
% \RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\usepackage{natbib}
\usepackage{adjustbox}
\usepackage{multirow}

@@

The figures herein are supplementary to those in the main manuscript. Figure 
\ref{fig:norm_dens} shows examples of marginal
posterior densities for QM of quantiles for a normal distribution from the 
simulation study in section 4.1 in
the main manuscript. The parameter uncertainty of the QGP and ORD models are 
unsurprisingly similar in
most cases, and where they differ it is hard to say that one is estimating the 
true parameter better than the
other. The IND model on the other hand has tighter densities, and often the 
bulk of the posterior is far from
the true parameter.

\begin{figure}
\centering
  \includegraphics[width = 1.1\linewidth]{Images/norm_params_fit.png}
\caption{Density plots of posterior distribution samples for normal parameters 
by QM for QGP, ORD, and
IND models. QM was done on estimated quantiles from a normal distribution with 
mean 4 and standard
deviation 3.5. The posterior densities are for $\mu$ (top left), 
$\sigma$ (top right), and 
sample size n (bottom). Plots are
faceted by the sample size $n \in \{50, 150, 500, 1,000\}$ 
($y$-axis) and number of 
quantiles $K \in \{7, 15, 23\}$ ($x$-axis).
Vertical lines (black) show the value of the true parameter}
\label{fig:norm_dens}
\end{figure}

Figures \ref{fig:exp_dens} and \ref{fig:exp_cov_dists} are from a simulation 
study similar to that of section 4.1 
except that QM was performed on
quantiles estimated from draws from an exponential distribution with 
parameter $\lambda = 4$ rather than a normal distribution. 
The results in the study for the exponential distribution were similar 
to those for the normal
distribution.

\begin{figure}
\centering
  \includegraphics[width = 1.1\linewidth]{Images/exp_params_fit.png}
\caption{Density plots of posterior distribution samples for the exponential 
parameters by QM for QGP,
ORD, and IND models. QM was done on estimated quantiles from a exponential 
distribution with parameter
$\lambda = 4$. The posterior densities are for $\lambda$
(left) and sample size $n$ (right). 
Plots are faceted by true sample
size $n \in \{50, 150, 500, 1,000\}$ ($y$-axis) and number of quantiles 
$K \in \{7, 15, 23\}$ 
($x$-axis). Vertical lines (black)
show the value of the true parameter.}
\label{fig:exp_dens}
\end{figure}

\begin{figure}[hbt!]
% \begin{subfigure}
  \includegraphics[width=\linewidth]{Images/exp_cov_dists.png}

\caption{ Posterior coverage (top) calculated as the percentage of times the 
true parameter fell within the
modeled 90\% credible interval over the 500 replications. Coverage is faceted 
by the exponential parameter $\lambda$
and $n$ with $K = 23$, and by increasing sample size ($x$-axis). 
The five models 
QGP, ORD, QGPN, ORDN,
and IND are colored as shown the legend. The horizontal line (black) is at 
the nominal 90\% level. Only QGP
and ORD appear for the parameter $n$ as they are the only two which estimate 
an unknown $n$. (bottom)
Distance between the true distribution and the estimated QM predictive 
distribution averaged over the 500
replications. Distances include UWD1, TV, and KLD for $K = 23$, 
and by increasing sample size (x-axis)}
\label{fig:exp_cov_dists}
\end{figure}

??Figures 11, 12, and 13 were referred to
in section 4.2 of the main manuscript.??

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1.3\linewidth]{Images/quants_dens_la.png}

\caption{QM fits of $K=23$ quantiles by KDE, SPL, IND, and QGP for $n \in \{50, 150, 500, 1{,}000\}$. The quantiles were sampled from the Laplace distribution $La(0,1)$. The quantile fits (left) show the true quantiles (black) with either the QM fit line (grey) or the credible intervals of 50\% (red) and 95\% (pink). 
The estimated PDF plots (right) show the true PDF (black) with either a the QM estimated PDF (grey) or the credible intervals of 50\% (red) and 95\% (pink).}
\label{fig:lp_fits}
\end{figure}

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1.3\linewidth]{Images/quants_dens_gmix.png}
\caption{QM fits of $K=23$ quantiles by KDE, SPL, IND, and QGP for $n \in \{50, 150, 500, 1{,}000\}$. The quantiles were sampled from the two component normal mixture distribution $w N(-1, 0.9) + (1-w)N(1.2, .6)$ where $w = 0.35$. The quantile fits (left) show the true quantiles (black) with either the QM fit line (grey) or the credible intervals of 50\% (red) and 95\% (pink). 
The estimated PDF plots (right) show the true PDF (black) with either a the QM estimated PDF (grey) or the credible intervals of 50\% (red) and 95\% (pink). }
\label{fig:gmix_fits}
\end{figure}

\begin{figure}[hbt!]
\centering
  \includegraphics[]{Images/mix_comps.png}
\caption{Future caption}
\label{fig:mix_comps}
\end{figure}

% \begin{table}[hbt!]
% \footnotesize
% \centering
% \caption{UWD1}
% \begin{adjustbox}{center}
% \begin{tabular}{lcccccc|cccccc}
%  & \multicolumn{6}{c}{QGP} & \multicolumn{6}{c}{QGP-n} \\
% \cmidrule(lr){2-7} \cmidrule(lr){8-13}
% Components & 1 & 2 & 3 & 4 & 5 & 6 & 1 & 2 & 3 & 4 & 5 & 6 \\
% \midrule
% EV & 0.063 & 0.025 & 0.021 & 0.020 & 0.020 & 0.020 & 
%       0.064 & 0.025 & 0.021 & 0.021 & 0.020 & 0.020\\
% La & 0.044 & 0.022 & 0.020 & 0.020 & 0.020 & 0.020 & 
%       0.044 & 0.021 & 0.020 & 0.020 & 0.020 & 0.020 \\
% MIX & 0.107 & 0.018 & 0.019 & 0.019 & 0.019 & 0.019 & 
%       0.157 & 0.018 & 0.019 & 0.018 & 0.018 & 0.018
% \end{tabular}
% \end{adjustbox}
% \end{table}

\begin{table}[hbt!]
\footnotesize
\centering
\caption{UWD1}
\begin{adjustbox}{center}
\begin{tabular}{clcccccc|cccccc}
& & \multicolumn{6}{c}{QGP} & \multicolumn{6}{c}{QGP-n} \\
\cmidrule(lr){3-8} \cmidrule(lr){9-14}
& Components & 1 & 2 & 3 & 4 & 5 & 6 & 1 & 2 & 3 & 4 & 5 & 6 \\
\midrule
\multirow{3}{*}{UWD1} &
EV & 0.063 & 0.025 & 0.021 & 0.020 & 0.020 & 0.020 & 
      0.064 & 0.025 & 0.021 & 0.021 & 0.020 & 0.020\\
 & La & 0.044 & 0.022 & 0.020 & 0.020 & 0.020 & 0.020 & 
      0.044 & 0.021 & 0.020 & 0.020 & 0.020 & 0.020 \\
 & MIX & 0.107 & 0.018 & 0.019 & 0.019 & 0.019 & 0.019 & 
      0.157 & 0.018 & 0.019 & 0.018 & 0.018 & 0.018 \\
      \hline
\multirow{3}{*}{TV} &
EV & 0.127 & 0.053 & 0.047 & 0.047 & 0.048 & 0.049 &
      0.125 & 0.052 & 0.039 & 0.039 & 0.039 & 0.039\\
 & La & 0.117 & 0.051 & 0.047 & 0.047 & 0.048 & 0.048 &
      0.117 & 0.048 & 0.043 & 0.043 & 0.043 & 0.042 \\
 & MIX & 0.226 & 0.038 & 0.033 & 0.032 & 0.033 & 0.034 &
      0.225 & 0.029 & 0.031 & 0.032 & 0.034 & 0.035 \\
      \hline
\multirow{3}{*}{KLD} &
EV & 0.171 & 0.025 & 0.014 & 0.013 & 0.014 & 0.015 & 
      0.170 & 0.025 & 0.011 & 0.010 & 0.010 & 0.010\\
& La & 0.110 & 0.014 & 0.012 & 0.012 & 0.012 & 0.012 & 
      0.111 & 0.013 & 0.010 & 0.010 & 0.010 & 0.010 \\
& MIX & 0.221 & 0.010 & 0.008 & 0.008 & 0.009 & 0.010 & 
      0.293 & 0.004 & 0.006 & 0.007 & 0.009 & 0.010
\end{tabular}
\end{adjustbox}
\end{table}

\end{document}

@@

% \begin{thebibliography}{00}
% 
% %% For authoryear reference style
% %% \bibitem[Author(year)]{label}
% %% Text of bibliographic item
% 
% \bibitem[Lamport(1994)]{lamport94}
%   Leslie Lamport,
%   \textit{\LaTeX: a document preparation system},
%   Addison Wesley, Massachusetts,
%   2nd edition,
%   1994.
% 
% \end{thebibliography}
